---
title: "Organized"
author: "Axel Leon"
date: "9/15/2021"
output:
  html_document:
    css: vintage.css
    fig_caption: yes
    highlight: pygments
#  theme: simplex   To have the TOC in Red
    toc: yes
    toc_float: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}

Sys.Date()
```

***

## Objective 

This Rmd is to learn how to format the data prior to analyses.  The data are collected form the Karn and Penrose, in the same structure they present in the paper, for analysis they needed to be wrangled for effective analysis in R and RStudio.



```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(gt)
library(reshape)
library(janitor)
```
***

A small replicable data set

```{r}
mini=tribble( ~mother_age,~non_survivors,~total_birth,~birth_weight,
14,0,1,3.5,
16,1,1,4,
17,0,1,4.5,
18,0,2,5,
19,0,4,6,
20,0,5,6.5,
22,3,5,7,
25,4,10,7.5)

mini

sum(mini$total_birth- mini$non_survivors)
```
***

Create a data frame for all death **0** and surviving babies **1**.

- Calculate the number of survivors
- Assign to a new column the code of **0** for all dead babies
- Assign to a new column the code of **0** for a=ll babies alive
- Untable (expand) the data set the number of times each conditions are present
  - mother's age
  - birth weight
- bind both data frames in one
- arrange (sort) by age of mother or weight (your preference)

```{r}
df=as.data.frame(mini)%>%
  mutate(survivors = total_birth - non_survivors)
df
sum(df$total_birth)
## untable non-survivors
df1<- untable(df[,c(1,2,4,5)], num = df[,"non_survivors"])%>%
  filter(non_survivors >0)%>%
  mutate(Alive_Dead = ifelse(non_survivors >0,0, NA))
df1


## untable survivors
df2<- untable(df[,c(1,2,5,4)], num = df[,"survivors"])%>%
  filter(survivors >0)%>%
  mutate(Alive_Dead = ifelse(survivors >0, 1, NA))
df2

#df3 <- merge(df1, df2) Not appropirate function
#df3 %>% arrange(mother_age)
#unique(df3$dead)

#df4=rbind(df1, df2)  #Function in base R
#df4 %>% arrange(mother_age)

df5=bind_rows(df1, df2) # Function in dplyr

df5 %>% arrange(mother_age)

row.names(df5)<-NULL  # this removes the row names
df5 %>% arrange(mother_age)
```
The above functions are now to be applied to the data sets of Karn and Penrose


***


```{r}
library(readr)
T2i <- read_csv("DATA/T2i.csv")
T2i

```








```{r}
df=as.data.frame(T2i)%>%
  mutate(survivors = Count_Total_Birth - Count_Non_survivors)
head(df, n=8)
df1<- untable(df[,c(1,2,4,5)], num = df[,3])%>%
  filter(Count_Non_survivors > 0)%>%
  mutate(dead = ifelse(Count_Non_survivors>=0,0))


df2<- untable(df[,c(1,2,5,4)], num = df[,2])%>%
  filter(survivors >0)%>%
  mutate(Survivors = ifelse(survivors >0, 1))
df3 <- merge(df1, df2)
head(df3, n=8)
unique(df3$dead)
```




```{r}
7.424e-4
```

